Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-02-14 abstract.tex
:sparkles: Refactor code

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/pc_power_sx.tex}
			}
			\caption{Average channel power versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$. $(N^\mathrm{T}, N^\mathrm{R}) = (8, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
			\label{sm:pc_power_sx}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\subfloat[Without Direct Link\label{sm:pc_power_bond_nd}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_power_bond_nd.tex}
				}
			}
			\subfloat[With Direct Link\label{sm:pc_power_bond_hd}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_power_bond_hd.tex}
				}
			}
			\caption{Average channel power versus \gls{ris} group size $L$. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (8, 256, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
			\label{sm:pc_power_bond}
		\end{figure}
		% Fig.~\ref{sm:pc_power_sx} shows that, apart from numerous reflecting elements, a sufficiently large group size can help
		% Fig.~\ref{sm:pc_power_sx} shows that increasing the group size can improve the channel power especially for a large \gls{ris}.
		Fig.~\ref{sm:pc_power_sx} shows that, apart from adding reflecting elements $N^\mathrm{S}$, increasing the group size $L$ also improves the channel power.
		% The behavior is more obvious for a large \gls{ris}.
		This behavior is more pronounced for a large \gls{ris}.
		For example, the gain of pairwise connection is \qty{2.8}{\percent} for $N^\mathrm{S} = 16$ and \qty{28}{\percent} for $N^\mathrm{S} = 256$.
		% \gls{bd} \gls{ris} provides a higher channel power than conventional \gls{ris}.
		% As $L$ increases from 1 to 2,
		It implies that the channel shaping capability of \gls{bd} \gls{ris} scales with group size $L$.

@@

	\begin{subsection}{Channel Singular Value Redistribution}
		We first show the channel shaping benefit of \gls{bd} \gls{ris} by a toy example.
		Consider $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 2, 2)$ and assume the direct link is absent.
		The diagonal \gls{ris} is $\mathbf{\Theta}^\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}^\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		When the direct link is absent, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		For a fair comparison, we enforce symmetry with $\beta = \pi / 2$.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/pc_singular_toy.eps}
			\caption{Channel singular value shaping by diagonal and symmetry unitary \gls{ris}. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 2, 2)$. Direct link is absent.}
			\label{sm:pc_singular_toy}
		\end{figure}
		Fig.~\ref{sm:pc_singular_toy} illustrates all possible channel singular values achieved by diagonal and symmetry unitary \gls{ris}.
		Despite using the same number of elements and parameters, \gls{bd} \gls{ris} provides much wider dynamic ranges of $\sigma_1(\mathbf{H})$ and $\sigma_2(\mathbf{H})$ than diagonal \gls{ris}.
		Larger gaps are expected when the symmetry constraint can be relaxed.

		We then analyze the channel shaping \emph{capability} of \gls{bd} \gls{ris} under specific setups.
		\begin{subsubsection}{Rank-Deficient Channel}
			\label{sc:pc_rank_deficient}
			In rank-deficient channels, \gls{bd} \gls{ris} $\mathbf{\Theta}^\mathrm{B}$ cannot achieve a higher \gls{dof} than diagonal \gls{ris} $\mathbf{\Theta}^\mathrm{D}$.
			This is because $\mathrm{sv}(\mathbf{\Theta}^\mathrm{B}) = \mathrm{sv}(\mathbf{\Theta}^\mathrm{D}) = \boldsymbol{1}$ and
			\begin{equation}
				\begin{split}
					\mathrm{rank}(\mathbf{H})
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \mathrm{rank}(\mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F}) \\
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \min \bigl( \mathrm{rank}(\mathbf{H}^\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}), \mathrm{rank}(\mathbf{H}^\mathrm{F}) \bigr).
				\end{split}
			\end{equation}
			% However, \gls{bd} \gls{ris} can still guarantee a higher rate at finite \gls{snr}.
			Note \gls{bd} \gls{ris} can still provide a higher indirect \gls{snr} as shown in Fig.~\ref{sm:pc_power_sx} and \ref{sm:pc_power_bond}.
		\end{subsubsection}

@@

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/pc_singular_pareto.tex}
			}
			\caption{Singular value Pareto front. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 64, 2)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
			\label{sm:pc_singular_pareto}
		\end{figure}

@@

		The Pareto front and evolving trend of channel singular values are shown in Fig.~\ref{sm:pc_singular_pareto} and \ref{sm:pc_singular_bound}.
		Clearly, \gls{bd} \gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}
\end{section}

>>>>

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/power_sx.tex}
			}
			\caption{Average channel power versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$. $(N^\mathrm{T}, N^\mathrm{R}) = (8, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
			\label{sm:power_sx}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\subfloat[Without Direct Link\label{sm:pc_power_bond_nd}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_power_bond_nd.tex}
				}
			}
			\subfloat[With Direct Link\label{sm:pc_power_bond_hd}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_power_bond_hd.tex}
				}
			}
			\caption{Average channel power versus \gls{ris} group size $L$. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (8, 256, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
			\label{sm:pc_power_bond}
		\end{figure}
		% Fig.~\ref{sm:power_sx} shows that, apart from numerous reflecting elements, a sufficiently large group size can help
		% Fig.~\ref{sm:power_sx} shows that increasing the group size can improve the channel power especially for a large \gls{ris}.
		Fig.~\ref{sm:power_sx} shows that, apart from adding reflecting elements $N^\mathrm{S}$, increasing the group size $L$ also improves the channel power.
		% The behavior is more obvious for a large \gls{ris}.
		This behavior is more pronounced for a large \gls{ris}.
		For example, the gain of pairwise connection is \qty{2.8}{\percent} for $N^\mathrm{S} = 16$ and \qty{28}{\percent} for $N^\mathrm{S} = 256$.
		% \gls{bd} \gls{ris} provides a higher channel power than conventional \gls{ris}.
		% As $L$ increases from 1 to 2,
		It implies that the channel shaping capability of \gls{bd} \gls{ris} scales with group size $L$.

@@

	\begin{subsection}{Channel Singular Value Redistribution}
		We first show the channel shaping benefit of \gls{bd} \gls{ris} by a toy example.
		Consider $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 2, 2)$ and assume the direct link is absent.
		The diagonal \gls{ris} is $\mathbf{\Theta}^\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}^\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		When the direct link is absent, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		For a fair comparison, we enforce symmetry with $\beta = \pi / 2$.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/singular_trend.eps}
			\caption{Channel singular value shaping by diagonal and symmetry unitary \gls{ris}. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 2, 2)$. Direct link is absent.}
			\label{sm:singular_trend}
		\end{figure}
		Fig.~\ref{sm:singular_trend} illustrates all possible channel singular values achieved by diagonal and symmetry unitary \gls{ris}.
		Despite using the same number of elements and parameters, \gls{bd} \gls{ris} provides much wider dynamic ranges of $\sigma_1(\mathbf{H})$ and $\sigma_2(\mathbf{H})$ than diagonal \gls{ris}.
		Larger gaps are expected when the symmetry constraint can be relaxed.

		We then analyze the channel shaping \emph{capability} of \gls{bd} \gls{ris} under specific setups.
		\begin{subsubsection}{Rank-Deficient Channel}
			\label{sc:pc_rank_deficient}
			In rank-deficient channels, \gls{bd} \gls{ris} $\mathbf{\Theta}^\mathrm{B}$ cannot achieve a higher \gls{dof} than diagonal \gls{ris} $\mathbf{\Theta}^\mathrm{D}$.
			This is because $\mathrm{sv}(\mathbf{\Theta}^\mathrm{B}) = \mathrm{sv}(\mathbf{\Theta}^\mathrm{D}) = \boldsymbol{1}$ and
			\begin{equation}
				\begin{split}
					\mathrm{rank}(\mathbf{H})
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \mathrm{rank}(\mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F}) \\
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \min \bigl( \mathrm{rank}(\mathbf{H}^\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}), \mathrm{rank}(\mathbf{H}^\mathrm{F}) \bigr).
				\end{split}
			\end{equation}
			% However, \gls{bd} \gls{ris} can still guarantee a higher rate at finite \gls{snr}.
			Note \gls{bd} \gls{ris} can still provide a higher indirect \gls{snr} as shown in Fig.~\ref{sm:power_sx} and \ref{sm:pc_power_bond}.
		\end{subsubsection}

@@

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/singular_pareto.tex}
			}
			\caption{Singular value Pareto front. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 64, 2)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
			\label{sm:singular_pareto}
		\end{figure}

@@

		The Pareto front and evolving trend of channel singular values are shown in Fig.~\ref{sm:singular_pareto} and \ref{sm:pc_singular_bound}.
		Clearly, \gls{bd} \gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}
\end{section}
